---
title: 'Sunspot Number : Explanations and Visualization of the algorithm to fill missing values'
author: "Antoine Pissoort"
date: "March 9, 2017"
output:
  html_document:  
    css: style.css
    collapsed: yes # Crée un document unique 
    fig_caption: yes # Figures encapsulées ? 
    fig_height: 4.5 # Hauteur par défaut des figures
    fig_width: 6.5 # Largeur par défaut des figure
    theme: lumen  # Style du document
    highlight: tango # style de mise en valeur du code
    code_folding: hide #  Cache le code  
    number_sections: yes # Ajout table des matières 
    toc: yes # Table des matière ?
    toc_depth: 3  # Profondeur table des matière
    toc_float: no # table des matière flottante
  word_document: default
#bibliography: P2consu.bib
---

```{r setup, include=T, message = F, warning=F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(microbenchmark)
library(profvis)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(reshape2)
library(plotly)
library(parallel)
library(pander)
setwd("H:/LinuxMint/PissoortRepo/ValUSunSSN/ValUSunSSN")
#source('smooth.gaussR_LOAD.R')  # Re-created Gaussian smoother
library(ValUSunSSN)
#load('/home/piss/PissoortRepo/ValUSunSSN/Data/DataSSN.RData')   # Take full dataset
#load('DataSSN81.RData')  # Take data since 1981
# data.chris <- read.csv("Alldata.csv")
# 
# data.chris$Date <- as.Date(paste(data.chris$year,
#                                  data.chris$month, data.chris$day,sep = "-"))
# data.chris[data.chris == -1] <- NA
# 
# data.chris <- data.chris[, colnames(data.chris) %in% c("Ng", "Ns") == F]
# # We do it only for SSN here
# 
# data.mat <- spread(data.chris, key="station", value="SSN")
# data.mat <- data.mat[order(data.mat$decdate),]
# 
# data.mat2 <- data.mat[,-(1:5)] # Get matrix of stations in columns only.
runExample()
```


# Explanations of the main Function 

All the functions used are available in the $\texttt{interpol2_em_pissoort.R}$. We will focus here on the main functions, in order to not being "overflowed" of code ...


```{r, eval = F}

```



# Run the algorithms  : Tests 

To do this, we will 
We ran the algorithm three times : 

* Once for the full dataset (since 1924) because we thought should not waste the opportunity to use data. 

* Since 1960 because it was the time period use by the $\texttt{FillInChris.R}$ from C.Ritter 

* Since 1981 as it seemed to be the most reliable time period regarding L.Lefèvre for example, and this also the time span used by "Véronique".

Moreover, comparisons between the results provided by the different time spans could be interesting. We could for example train a model using data from 1981 and then validate it using previous period, in order to have an independent measure of accuracy (...) 

and once in for the unmodified SSN datase
We take 

```{r, fig.height=7, eval =F}

## First, we discard stations that have coverage <5% for reliability issues

# If keep these values, the results will have incorrect values...
notNa.sum <- apply(data.mat2, 2, function(x) sum(!is.na(x)))
is_less5_cov <- unname(notNa.sum) < .05 * nrow(data.mat2)
data.mat2.fin <- data.mat2[, !is_less5_cov] # 7 stations are discarded !! 

#data.mat2.fin <- data.mat2.fin[data.mat$decdate >= 1981, ] # take only from 1981 ?


# Take this for input, as advised in the test.m file 
y <- sqrt(data.mat2.fin+1) # Select randomly here, for testing

options(mc.cores=parallel::detectCores()) # all available cores

z <- interpolsvd_em(y, nembed = 2, nsmo = 81, ncomp = 4,
                      niter = 30, displ = F)
# 193 sec for the whole dataset (with some stations discarded)
z <- z$y.filled


zssn <- interpolsvd_em(data.mat2.fin, nembed = 2, nsmo = 81, ncomp = 4,
                        niter = 30, displ = F)
zssn <- zssn$y.filled

```



```{r, eval=T}
rownames(y) <- rownames(zssn) <- 1:nrow(zssn)
y1 <- cbind(as.data.frame(data.mat2.fin), Date = data.mat$Date, x = "Raw")

z1 <- cbind(as.data.frame(zssn), Date = data.mat$Date[data.mat$decdate>1981], x = "Filled")
colnames(y1) <- colnames(z1) <- c(colnames(data.mat2.fin), "Date", "x")

zz <- rbind(y1,z1)

```


## Comparisons with raw data

First of all, we decide to show only the **station of Uccle** which is considered to some extent, as the "*main* station". In this dynamic plot provided by $\texttt{plotly}$, you can hide the class you would like to see, e.g. to see raw data you can click on both red and blue button. This plot can conveniently show us, among other things, how the filled alues "fits" the shape of the raw data. 

```{r, fig.width=9.5}
zz_neg_ucc <- data.frame( wnUC2 = zz$wnUC2, Date = zz$Date, x = zz$x, negative = zz$wnUC2 < 0 )
#z1$x <- ifelse(z1$wnUC2 < 0, "Filled <0", "Filled" ) 

#zz_neg_ucc[!is.na(zz_neg_ucc$x == "Filled"), ][,"x"] <- NA
## Visualize "incorrect" filled values, ie those below 0 
# zz_neg_ucc[!is.na(zz_neg_ucc$x == "Filled"), ][ ,"x"] <- ifelse(zz_neg_ucc[!is.na(zz_neg_ucc$x != "Raw"), ][ ,"wnUC2"] < 0, "Filled <0", "Filled" ) 
zz_neg_ucc[(nrow(zz_neg_ucc)/2+1):nrow(zz_neg_ucc), ][ ,"x"] <- ifelse(zz_neg_ucc[(nrow(zz_neg_ucc)/2+1):nrow(zz_neg_ucc), ][ ,"wnUC2"] < 0, as.factor("Filled <0"), as.factor("Filled") ) 
zz_neg_ucc[is.na(zz_neg_ucc$x),][,'x'] <- "Filled <0"
# zz_neg_ucc[zz_neg_ucc$x == "Filled", ]$x  <- ifelse( zz_neg_ucc[zz_neg_ucc$x == "Filled", ]$wnUC2 < 0, "Filled < 0", "Filled" )


# for (i in 1:(nrow(zz)/2) ){ 
#   if ( zz_neg_ucc[zz_neg_ucc$x == "Filled", ]$wnUC2[i] < 0 )
#     zz_neg_ucc[zz_neg_ucc$x == "Filled", ]$x[i] <- "Filled <0"
#   
#  # else  zz_neg_ucc[zz_neg_ucc$x == "Filled", ]$x[i] <- "Filled" 
# }


f <- list(
  family = "Calibri",
  size = 16,
  color = "#33666C"
)
f2 <- list(
  family = "Old Standard TT, serif",
  size = 14,
  color = "black"
)
xl <- list(
  title = " Date",
  titlefont = f,
  tickfont = f2,
   showticklabels = TRUE
)
yl <- list(
  title = "SSN for wnUC2",
  titlefont = f,
  tickfont = f2,
  showticklabels = TRUE
)
tit <- list(
  title = "Dynamic plot of the SSN for  the station of Uccle",
  titlefont = f
)
l <- list(
  font = list(
    family = "sans-serif",
    size = 13,
    color = "#000"),
  bgcolor = "#E2E2E2",
  bordercolor = "#FFFFFF",
  borderwidth = 2)

plot_ly(data = zz_neg_ucc, x = ~Date, y = ~`wnUC2`, type = "scattergl", color = ~x, colors = c("green", "blue", "red")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l)# %>% add_data(plot_ly(data = zz_neg_ucc[zz_neg_ucc$negative.wnUC2,], x = ~Date, y = ~`wnUC2`, type = "scattergl", colors = c("green"))) #%>% add_trace()%>% add_markers(color = ~as.factor(negative.wnUC2)) 

```

<br>

$\underline{\text{Comments}}$ : 

* We see that there seem to be some problems concerning the positivity of the values.

* 


For convenience, we decide to show now the two retained stations that has the greatest number of available values, i.e. 21465 (%) and 20855 (%) for *wnKS2* and *wnKZ2* respectively, and also the two stations with the lowest number of avialable values

```{r}
notNa.sum <- apply(data.mat2.fin, 2, function(x) sum(!is.na(x)))
#sort(notNa.sum)/nrow(data.mat2.fin) * 100

df <- data.frame(wnKS2 = "21465 (64.4%)", wnKZ2 = "18957 (56.9%) ", `wnFR-S` = "2119 (6.4%)", `wnGU-S` = "2185 (6.6%)" )
rownames(df) <- c("# available values ")
pander(df)
```


```{r, fig.width=10, fig.height=4}
yl$title <- "SSN for wnFR-S"
p1 <- plot_ly(data = zz, x = ~Date, y = ~`wnFR-S`, type = "scattergl", color = ~x, colors = c("red", "blue")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l)

yl$title <- "SSN for wnGU-S"
p2 <- plot_ly(data = zz, x = ~Date, y = ~`wnGU-S` , type = "scattergl", color = ~x, colors = c("red", "blue")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l) 

yl$title <- "SSN for wnKS2"
p3 <- plot_ly(data = zz, x = ~Date, y = ~wnKS2, type = "scattergl", color = ~x, colors = c("red", "blue")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l) 

yl$title <- "SSN for wnKZ2"
p4 <- plot_ly(data = zz, x = ~Date, y = ~wnKZ2, type = "scattergl", color = ~x, colors = c("red", "blue")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l) 

subplot(p1, p2, nrows = 1,shareY = T, titleY = T )
```

For these two stations, we remark that the small number of available values indeed impact the uncertainty and the way these missing values are filled by the algorithm. 

In contrast, when the number of available values is high $\Rightarrow$

```{r, fig.width=10, fig.height=4}
subplot(p3, p4, nrows = 1,shareY = T, titleY = T)
```

The filling seem quite more homogenous and thus robust.





```{r, fig.height=8, fig.width=9, eval=F}
# z1 <- cbind(as.data.frame(zz[nrow(data.mat),]), Date = data.mat$Date )
# colnames(z1) <- c(colnames(data.mat2.fin), "Date")
# zz.full <- melt(z1, id.vars = "Date" )
# dimnames(zz.full)[[2]][2] <- "Station"
# dimnames(zz.full)[[2]][3] <- "SSN"

g <- ggplot(zz.full) + geom_line(aes( x = Date, y = SSN, col = Station), size = .4) +
 scale_color_hue(l = 40, c = 200) + theme_piss()
  # scale_colour_brewer(name=expression(underline(bold("data: "))),
  #                     palette = "Set1") + 
  #guides(colour = guide_legend(override.aes = list(size= 2.5))) 

# ggplotly(g)

plot_ly(data = zz.full, x = ~Date, y = ~SSN, type = "scattergl", color = ~Station )

```


We can first notice a "problem", that is the values for SSN which are below 0

Could'nt we 




## Comparisons with other filling methods

### Filling method of "Vero", time series starting from 1981

**1. Are the non-missing values still the same ?**

We hope so, but we will check it.


```{r}
filledt.vero <- read.csv("/home/piss/PissoortRepo/ValUSunSSN/data/time_SSNf_1981.csv")

# By looking at the Julian dates at the top and the bottom of the series, we retrieved the dates.
filledt.vero$Date <- seq(as.Date("1981-02-19"), as.Date("2015-03-30"), by = 1)

#load('/home/piss/PissoortRepo/ValUSunSSN/Data/DataSSN81.RData')

colnames(zssn) <- colnames(z1[, -((ncol(z1)-1):(ncol(z1)))])
zssn <- cbind.data.frame(zssn, times = z1$Date)

zssn <- zssn[as.Date(zssn$times) %in% filledt.vero$Date,]

## add indicators for missingness
missings <- is.na()
colnames(missings) <- paste0("miss_", colnames(missings))
zssn_complete <- cbind.data.frame(zssn, missings)
sum(zssn_complete$miss_wnUC2)

head(dif <- zssn_complete$wnUC2[!zssn_complete$miss_wnUC2] -
       data.mat2.fin$wnUC2[!zssn_complete$miss_wnUC2] )
sum (dif) / 0.0024342

```


The difference actually the same for all the observations. this not really an intrinsic difference but rather a variable type difference; the first being integer and the latter being real (numeric).

**2. Comparisons with all data (missing & non-missing)**


```{r}
# df_uccle_vero <- data.frame(time = zssn$times[1:nrow(filledt.vero)], wnUC2 = filledt.vero$UC2, from = "Vero")
# 
# df_uccle_v <- rbind.data.frame(df_uccle_chris, data.frame(time = zssn$times[1:nrow(filledt.vero)], wnUC2 = zssn$wnUC2[1:nrow(filledt.vero)], from = "`interpolEM`"))
# 
# plot_ly(data = df_uccle_v, x = ~time, y = ~`wnUC2`, type = "scattergl", color = ~from, colors = c("green", "blue")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l)


df_uccle_vero <- data.frame(time = zssn$times, SSN_wnUC2 = filledt.vero$UC2, from = "Vero")

df_uccle_c <- rbind.data.frame(df_uccle_vero, data.frame(time = zssn$times, SSN_wnUC2 = zssn[,"wnUC2"], from = "`interpolEM`"))

ggplot(df_uccle_c, aes(col = from)) + geom_point(aes(x = time, y = SSN_wnUC2), size = 0.15) + 
  solar.cycle() + theme_piss() +  guides(colour = guide_legend(override.aes = list(size= 3))) + ggtitle("Filled missing values")
```


**Analyze residuals**


```{r}
resdf_vero <- data.frame(time = zssn$times,
                         res = (filledt.vero$UC2) - (zssn[,"wnUC2"]) ) 
max(resdf_vero$res)  ; min(resdf_vero$res)

library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(4, "Spectral")))

ggplot(resdf_vero, aes(x = time, y = res, col = res)) + geom_point( size = 0.35) + theme_piss() + solar.cycle() + scale_colour_gradientn(colours = myPalette(10), limits=c(-150, 250))
```



Or we could also visualize the discrepancies between the filled values

**Only for the filled data**

```{r}
filled_ssn <- zssn[,"wnUC2"][zssn_complete$miss_wnUC2]

resdf_vero <- data.frame(time = zssn$times[zssn_complete$miss_wnUC2],
                         res = (filledt.vero$UC2[zssn_complete$miss_wnUC2]) - filled_ssn ) 
max(resdf_vero$res)  ; min(resdf_vero$res)

myPalette <- colorRampPalette(rev(brewer.pal(4, "Spectral")))

# ggplot(resdf_vero, aes(x = time, y = res, col = res)) + geom_point( size = 0.35) + theme_piss() + solar.cycle() + scale_colour_gradientn(colours = myPalette(10), limits=c(-150, 250))
ggplot(resdf_vero, aes(col = res)) + geom_line(aes(x = time, y = res), size = 0.35)  + theme_piss() + solar.cycle() +
  scale_colour_gradientn(colours = myPalette(10), limits=c(-150, 250))
```



**Filling method of *C.Ritter*, time series starting from 1960**

Again, we will take th station of Uccle for reference.

```{r}
# filledt.chris <- read.csv("SSNfilledChris.csv") # Load data
# 
# load("DataSSN60.RData") # Load data we have build
# 
# filled.chris2 <- filledt.chris[,-2]
# 
# df_uccle_chris <- data.frame(time = filled.chris2$times, wnUC2 = filled.chris2$wnUC2_d.txt, from = "Chris")
# 
# df_uccle_c <- rbind.data.frame(df_uccle_chris, data.frame(time = filled.chris2$times, wnUC2 = zssn$wnUC2[1:nrow(df_uccle_chris)], from = "`interpolEM`"))
# 
# plot_ly(data = df_uccle_c, x = ~time, y = ~`wnUC2`, type = "scattergl", color = ~from, colors = c("green", "blue")) %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l)

filledt.chris <- read.csv("SSNfilledChris.csv") # Load data

load("DataSSN60.RData") # Load data we have build

filled.chris2 <- filledt.chris[,-2]

## Are the non-missing values still the same ?? 
## add indicators for missingness
missings <- is.na(y)
colnames(missings) <- paste0("miss_", colnames(missings))
zssn_complete <- cbind.data.frame(zssn, missings)
sum(zssn_complete$miss_wnUC2)


head(dif <- zssn_complete$wnUC2[!zssn_complete$miss_wnUC2] - data.mat2.fin$wnUC2[!zssn_complete$miss_wnUC2] )
sum (dif) / 0.0024342
```

Same idea. The very slight discrepancies are du to (type) approximations

```{r}
df_uccle_chris <- data.frame(time = filled.chris2$times, wnUC2 = filled.chris2$wnUC2_d.txt, from = "Chris")

df_uccle_c <- rbind.data.frame(df_uccle_chris, data.frame(time = filled.chris2$times, wnUC2 = zssn$wnUC2[1:nrow(df_uccle_chris)], from = "`interpolEM`"))


# Comparisons for first 5 years (1980 : 1985)
plot_ly(data = df_uccle_c[1980 < df_uccle_c$time & df_uccle_c$time < 1985 ,], x = ~time, y = ~`wnUC2`, color = ~from, type = "scattergl", colors = c("green", "blue"), mode =  "markers+text") %>% layout(xaxis = xl, yaxis = yl, title = tit, legend = l)


## Residuals

resdf_chris <- data.frame(time = z1$Date[1:nrow(filled.chris2)],
                         res = (filled.chris2$wnUC2_d.txt) - (zssn[,"wnUC2"][1:nrow(filled.chris2)]) ) 
max(resdf_chris$res)  ; min(resdf_chris$res)

# ggplot(resdf_chris, aes(x = time, y = res, col = res)) + geom_point( size = 0.35) + theme_piss() + solar.cycle() + scale_colour_gradientn(colours = myPalette(10), limits=c(-95, 88))
ggplot(resdf_chris, aes(col = res)) + geom_line(aes(x = time, y = res), size = 0.35)  + theme_piss() + solar.cycle() +
  scale_colour_gradientn(colours = myPalette(10),limits=c(-95, 88))
```


**Only for filled values :**

```{r}
filled_ssn <- zssn[,"wnUC2"][1:nrow(df_uccle_chris)][zssn_complete$miss_wnUC2]

resdf_chris <- data.frame(time = z1$Date[1:nrow(filled.chris2)][zssn_complete$miss_wnUC2],
                         res = (filled.chris2$wnUC2_d.txt[zssn_complete$miss_wnUC2]) - filled_ssn ) 
max(resdf_chris$res)  ; min(resdf_chris$res)

myPalette <- colorRampPalette(rev(brewer.pal(4, "Spectral")))

# ggplot(resdf_chris, aes(x = time, y = res, col = res)) + geom_point( size = 0.35) + theme_piss() + solar.cycle() + scale_colour_gradientn(colours = myPalette(10), limits=c(-150, 250))
ggplot(resdf_chris, aes(col = res)) + geom_line(aes(x = time, y = res), size = 0.35)  + theme_piss() + solar.cycle() +
  scale_colour_gradientn(colours = myPalette(10), limits=c(-150, 250))

```

Discrepancies are less than with Vero’s method…

We could want want to pick all the time or station that have big discrepancies. This could be informative






```{r}
library(softImpute)
vignette(package = "softImpute")

data.mat2 <- apply(data.mat2, 2, as.numeric)

fit <- softImpute(data.mat2, trace = T, rank.max = 20, lambda = 2, type = "svd")
fit <- softImpute(data.mat2, trace = T, rank.max = 20, lambda = 2, type = "als")

data.mat2.fil1 <- softImpute::complete(data.mat2, fit)


xs=as(data.mat2,"Incomplete")
xsc <- biScale(data.mat2, col.scale=FALSE,row.scale=FALSE)
```





```{r, eval=F}
# rownames(zssn) <- data.mat$decdate
# zssn <- as.data.frame(zssn) 
# colnames(zssn) <- colnames(data.mat2.fin) 
# write.csv(zssn, "SSNfillFULL_R.csv")
```




